Determining Causes of Death with ChatGPT: A Case Study of Deaths from Verbal Autopsy in Sierra Leone from 2019-2022

### 2023-12-15 manuscript/manuscript.tex
Working on bg

Most low-income countries have reported to have either no data or have registered less than 50\% of deaths in their country, where only 8\% of these deaths have a Cause of Death (COD) recorded \cite{worldhealthorganizationSCOREHealthData2021}. To increase the number of records with a COD, a method known as Verbal Autopsy (VA) is used to determine likely causes for deaths outside of healthcare facilities. VA involves trained interviewers that speak to those familiar with the deceased to gather information based on standard questionnaires and open narratives, while medical professionals evaluate the information to assign probable CODs \cite{worldhealthorganizationVerbalAutopsyStandards2023}.

>>>>

Accurate counts of deaths and diagnosis of their causes are crucial for public health planning and policy making --- guiding interventions to reduce premature deaths, and ensuring resources are allocated efficiently. However, most low-income countries have reported to have either no data on deaths or have registered less than 50\% of deaths in their country, where only 8\% of these deaths have a Cause of Death (COD) recorded \cite{worldhealthorganizationSCOREHealthData2021}. To increase the number of records with a COD, a method known as Verbal Autopsy (VA) is used to determine likely causes for deaths outside of healthcare facilities. VA involves trained interviewers that speak to those familiar with the deceased to gather information based on standard questionnaires and open narratives, while medical professionals evaluate the information to assign probable CODs \cite{worldhealthorganizationVerbalAutopsyStandards2023}.
